State of The Practice in Indirect Conflicts: Work Flows, Tools, and Future Direction

### 2014-03-10 iced-2013/paper.tex
ICED quick edits.

\begin{itemize}
	\item These false positives are caused by a difference in what
				developers consider to be false positives versus what awareness literature considers they are.
\end{itemize}

@@

\begin{itemize}
	\item Developers don't care about what they don't need to account for
\end{itemize}

@@

This unwanted awareness, or limited awareness, seems counter intuitive to current awareness understanding which would state that
being aware of all events in ones surrounding leads to higher productivity or other quality aspects. In theory this is correct,
but as it was found in practice, this is an incorrect assumption.
In regards to this full awareness, one developer said ``There is no room for this in [our company], as tools are already in place for
analysis of change[s] and code review takes care of the rest''. Since software developers have limits on their time, awareness of
all events surrounding a developer's work or project is not possible.
In terms of developer communication, Hattori et. al.~\cite{Hattori:2012:ICG} found through qualitative user studies, that developers
tend to use the bare amount of communication
towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose.
\textit{Developers prefer to spend their limited time dealing with
the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
of events occurring around their work which pose no direct threat to the consumption of their time.}

@@

\begin{itemize}
	\item Through unwanted awareness, and the use of developer time,
				developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
				may never arise.
\end{itemize}

@@

The lack of prevention process and tools being used is believed to be due to 2 factors as discovered in the interview process. 
The first being the identification of dependencies.
Even with an experienced system architect, identifying dependencies and notifying those involved is a daunting task
which is ignored more than dealt with. The second being the knowledge of when a dependency will fail, also requires vast knowledge
of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts.
Hattori et. al.~\cite{Hattori:2012:ICG} show that with direct conflicts, the
sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges,
which is the situation many indirect conflict tools have strived for, however the 2 factors of dependencies and knowledge of failure
are what ultimately have led to the amount of false positives and information overload seen in previous tools. The
abundance of detection and curative process and tools on the other hand shows once again developers willingness to debug
real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. The question ultimately
presented in this area is if curative measures are more productive than preventative measures.

@@

A deeper understanding of what context indirect conflicts occur in seems to be more of a success factor to indirect conflict
research than may have been previously thought. Static analysis may be useful in regards to indirect conflict context in that we
could identify trends surrounding the mature points of previous projects in order to give a better understanding of what it
means for a project to be pre or post mature point, which could affect the outcome of indirect conflict understanding.
Fluri et al.~\cite{Fluri:2007:CDT} uses static analysis and abstract syntax trees to be able to determine what types of source code
changes occur at a given point in time which could be used for determining such a context of mature points.

@@

Aside from context, dependency identification and tracking is a key missing component of
indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and
organizational structures of software teams. Between increased modularization (multiple sub projects or repositories),
cross language dependencies, and languages which do not lend themselves to static analysis~\cite{Jensen:2011:MHD}, static analysis tools have
become quite limited in identifying and tracking dependencies where they were once strong.
In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards the
ill fit of current static analysis to current development practices..

>>>>

\textbf{These false positives are caused by a difference in what
developers consider to be false positives versus what awareness literature considers they are.}

@@

\textbf{Developers don't care about what they don't need to account for}

@@

This unwanted awareness, or limited awareness, seems counter intuitive to current awareness understanding which would state that
being aware of all events in ones surrounding leads to higher productivity or other quality aspects. In theory this is correct,
but as it was found in practice, this is an incorrect assumption.
In regards to this full awareness, one developer said ``There is no room for this in [our company], as tools are already in place for
analysis of change[s] and code review takes care of the rest''. Since software developers have limits on their time, awareness of
all events surrounding a developer's work or project is not possible.
In terms of developer communication, Hattori et. al.~\cite{Hattori:2012:ICG} found through qualitative user studies, that developers
tend to use the bare amount of communication
towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose.
\textbf{Developers prefer to spend their limited time dealing with
the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
of events occurring around their work which pose no direct threat to the consumption of their time.}

@@

\textbf{Through unwanted awareness, and the use of developer time,
				developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
				may never arise.}

@@

The lack of prevention process and tools being used is believed to be due to 2 factors as discovered in the interview process. 
The first being the identification of dependencies.
Even with an experienced system architect, identifying dependencies and notifying those involved is a daunting task
which is ignored more than dealt with. The second being the knowledge of when a dependency will fail, also requires vast knowledge
of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts.
Hattori et. al.~\cite{Hattori:2012:ICG} show that with direct conflicts, the
sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges,
which is the situation many indirect conflict tools have strived for, however the 2 factors of dependencies and knowledge of failure
are what ultimately have led to the amount of false positives and information overload seen in previous tools. \textbf{The
abundance of detection and curative process and tools shows once again developers willingness to debug
real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. The question ultimately
presented in this area is if curative measures are more productive than preventative measures.}

@@

\textbf{A deeper understanding of what context indirect conflicts occur in seems to be more of a success factor to indirect conflict
research than may have been previously thought.} Static analysis may be useful in regards to indirect conflict context in that we
could identify trends surrounding the mature points of previous projects in order to give a better understanding of what it
means for a project to be pre or post mature point, which could affect the outcome of indirect conflict understanding.
Fluri et al.~\cite{Fluri:2007:CDT} uses static analysis and abstract syntax trees to be able to determine what types of source code
changes occur at a given point in time which could be used for determining such a context of mature points.

@@

Aside from context, dependency identification and tracking is a key missing component of
indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and
organizational structures of software teams. \textbf{Between increased modularization (multiple sub projects or repositories),
cross language dependencies, and languages which do not lend themselves to static analysis~\cite{Jensen:2011:MHD}, static analysis tools have
become quite limited in identifying and tracking dependencies where they were once strong.}
In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards the
ill fit of current static analysis to current development practices..
